
{{< include _math.qmd >}}

```{r setup, include=FALSE}
source('_setup.R')
conflict_prefer('lag', 'dplyr')
conflict_prefer('lead', 'dplyr')
```


# Temporadas 2014 a 2023

## Leitura

Para cada jogador, a coluna `all_nba` diz se ele foi para a *expert list* ou não.

A *expert list* está sendo considerada um conjunto, sem ordem.

```{r}
df <- read_csv(
  'dados/j-years-with-all-nba.csv'
) %>% 
  clean_names()
```


```{r}
df %>% dfSummary() %>% print(method = 'render')
```

::: {.callout-caution title="Colunas normalizadas"}

Todas as colunas numéricas estão normalizadas entre $0$ e $1$.

:::


## Colunas

De <https://www.basketball-reference.com/about/glossary.html>, em ordem alfabética. 

Ver também <https://www.nba.com/stats/help/glossary>.

::: {.callout-warning title="Faltaram"}

* dbpm: defensive bpm???
* f_tr: ???
* obpm: offensive bpm???
* x3p_ar: ???

:::

2P (x2p)
: 2-Point Field Goals

2P% (x2p_percent)
: 2-Point Field Goal Percentage; the formula is $2P/2PA$.

2PA (x2pa)
: 2-Point Field Goal Attempts

3P (x3p)
: 3-Point Field Goals (available since the 1979-80 season in the NBA)

3P% (x3p_percent)
: 3-Point Field Goal Percentage (available since the 1979-80 season in the NBA); the formula is $3P / 3PA$.

3PA (x3pa)
: 3-Point Field Goal Attempts (available since the 1979-80 season in the NBA)

Age
: Age; player age on February 1 of the given season.

AST
: Assists

AST%
: Assist Percentage (available since the 1964-65 season in the NBA); the formula is $100 * AST / (((MP / (Tm MP / 5)) * Tm FG) - FG)$. Assist percentage is an estimate of the percentage of teammate field goals a player assisted while he was on the floor.

BLK
: Blocks (available since the 1973-74 season in the NBA)

BLK%
: Block Percentage (available since the 1973-74 season in the NBA); the formula is $100 * (BLK * (Tm MP / 5)) / (MP * (Opp FGA - Opp 3PA))$. Block percentage is an estimate of the percentage of opponent two-point field goal attempts blocked by the player while he was on the floor.

BPM
: Box Plus/Minus (available since the 1973-74 season in the NBA); a box score estimate of the points per 100 possessions that a player contributed above a league-average player, translated to an average team. Please see the article About Box Plus/Minus (BPM) for more information.

DRB
: Defensive Rebounds (available since the 1973-74 season in the NBA)

DRB%
: Defensive Rebound Percentage (available since the 1970-71 season in the NBA); the formula is $100 * (DRB * (Tm MP / 5)) / (MP * (Tm DRB + Opp ORB))$. Defensive rebound percentage is an estimate of the percentage of available defensive rebounds a player grabbed while he was on the floor.

DRtg (???)
: Defensive Rating (available since the 1973-74 season in the NBA); for players and teams it is points allowed per 100 posessions. This rating was developed by Dean Oliver, author of Basketball on Paper. Please see the article Calculating Individual Offensive and Defensive Ratings for more information.

DWS
: Defensive Win Shares; please see the article Calculating Win Shares for more information.

eFG%
: Effective Field Goal Percentage; the formula is $(FG + 0.5 * 3P) / FGA$. This statistic adjusts for the fact that a 3-point field goal is worth one more point than a 2-point field goal. For example, suppose Player A goes 4 for 10 with 2 threes, while Player B goes 5 for 10 with 0 threes. Each player would have 10 points from field goals, and thus would have the same effective field goal percentage (50%).

FG
: Field Goals (includes both 2-point field goals and 3-point field goals)

FG%
: Field Goal Percentage; the formula is $FG / FGA$.

FGA
: Field Goal Attempts (includes both 2-point field goal attempts and 3-point field goal attempts)

FT
: Free Throws

FT%
: Free Throw Percentage; the formula is $FT / FTA$.

FTA
: Free Throw Attempts

G
: Games

GS
: Games Started (available since the 1982 season)

MP
: Minutes Played (available since the 1951-52 season)

ORtg (???)
: Offensive Rating (available since the 1977-78 season in the NBA); for players it is points produced per 100 posessions, while for teams it is points scored per 100 possessions. This rating was developed by Dean Oliver, author of Basketball on Paper. Please see the article Calculating Individual Offensive and Defensive Ratings for more information.

ORB
: Offensive Rebounds (available since the 1973-74 season in the NBA)

ORB%
: Offensive Rebound Percentage (available since the 1970-71 season in the NBA); the formula is $100 * (ORB * (Tm MP / 5)) / (MP * (Tm ORB + Opp DRB))$. Offensive rebound percentage is an estimate of the percentage of available offensive rebounds a player grabbed while he was on the floor.

OWS
: Offensive Win Shares; please see the article Calculating Win Shares for more information.

PER
: Player Efficiency Rating (available since the 1951-52 season); PER is a rating developed by ESPN.com columnist John Hollinger. In John's words, "The PER sums up all a player's positive accomplishments, subtracts the negative accomplishments, and returns a per-minute rating of a player's performance." Please see the article Calculating PER for more information.

PF
: Personal Fouls

PTS
: Points

STL
: Steals (available since the 1973-74 season in the NBA)

STL%
: Steal Percentage (available since the 1973-74 season in the NBA); the formula is $100 * (STL * (Tm MP / 5)) / (MP * Opp Poss)$. Steal Percentage is an estimate of the percentage of opponent possessions that end with a steal by the player while he was on the floor.

Tm
: Team

TOV
: Turnovers (available since the 1977-78 season in the NBA)

TOV%
: Turnover Percentage (available since the 1977-78 season in the NBA); the formula is $100 * TOV / (FGA + 0.44 * FTA + TOV)$. Turnover percentage is an estimate of turnovers per 100 plays.

TRB
: Total Rebounds (available since the 1950-51 season)

TRB%
: Total Rebound Percentage (available since the 1970-71 season in the NBA); the formula is $100 * (TRB * (Tm MP / 5)) / (MP * (Tm TRB + Opp TRB))$. Total rebound percentage is an estimate of the percentage of available rebounds a player grabbed while he was on the floor.

TS%
: True Shooting Percentage; the formula is $PTS / (2 * TSA)$. True shooting percentage is a measure of shooting efficiency that takes into account field goals, 3-point field goals, and free throws.

Usg%
: Usage Percentage (available since the 1977-78 season in the NBA); the formula is $100 * ((FGA + 0.44 * FTA + TOV) * (Tm MP / 5)) / (MP * (Tm FGA + 0.44 * Tm FTA + Tm TOV))$. Usage percentage is an estimate of the percentage of team plays used by a player while he was on the floor.

VORP
: Value Over Replacement Player (available since the 1973-74 season in the NBA); a box score estimate of the points per 100 TEAM possessions that a player contributed above a replacement-level (-2.0) player, translated to an average team and prorated to an 82-game season. Multiply by 2.70 to convert to wins over replacement. Please see the article About Box Plus/Minus (BPM) for more information.

WS
: Win Shares; an estimate of the number of wins contributed by a player. Please see the article Calculating Win Shares for more information.

WS/48 (ws_48)
: Win Shares Per 48 Minutes (available since the 1951-52 season in the NBA); an estimate of the number of wins contributed by the player per 48 minutes (league average is approximately 0.100). Please see the article Calculating Win Shares for more information.

::: {.callout-warning title="De onde vem defensive rating?"}

drtg é usada no DWARFF-BB, mas não está no arquivo csv.

:::


## Gravar csv

```{r}
df %>% 
  write_csv('dados/temporadas-2014-2023.csv')
```

